\usefont{OT1}{bch}{b}{n} \normalfont \normalsize \textsc{STEM Fellowship Big Data Challenge 2020} \\ [10pt] \huge Deep learning transcriptomic model for prediction of pan-drug chemotherapeutic sensitivity \\

### 2020-06-28 manuscript/manuscript.tex
updated flow chart

% Materials and Methods
\section{Methods}
We developed a deep learning model to accurately classify cancer cell lines into therapeutic response groups using data from the Genomics of Drug Sensitivity in Cancer (GDSC) database \cite{gdsc}. Following data collection and curation, we utilized unsupervised clustering algorithms to define two groups based on chemotherapeutic response. Next, we employed a biologically agnostic feature selection algorithm, Boruta, to select statistically relevant genes for our neural network. These transcriptomic features were fed into a neural network that classifies patients into therapeutic response groups as defined by the clustering algorithms. The neural network architecture was optimized using a grid search and evaluated using 5-fold stratified cross-validation. See Fig. \ref{fig:pipeline} for an overview of the data analysis pipeline. The source code can be found here: \url{https://github.com/tig3r66/youreka_genes}.

@@

% flowchart figure
\begin{figure}[t]
	\centering
	\begin{tikzpicture}[node distance=6em, auto]
	% blocks
	\node [block, text width=20em] (init) {Obtain data from the GDSC consortium (1,110 cell lines, 251 drugs).};
	\node [block, below of=init, text width=20em, yshift=-1em] (process) {Filter data: retain drugs with $>$80\% data for cell lines (548 cell lines, 117 drugs). Matrix normalization on IC50 via z-score.};
	\node [block, below of=process, text width=20em, yshift=-1em] (kmeans) {Unsupervised clustering into 2 chemotherapeutic response groups.};
	\node [block, below of=kmeans, text width=20em, yshift=-1em] (append) {Appended GDSC Affymetrix Human Genome U219 expression data to the IC50 matrix.};
	\node [block, below of=append, text width=20em, yshift=-1.5em] (feat) {Feature selection for relevant genes in RNA microarray data via Boruta algorithm. 300 genes selected at $\alpha=0.05$.};
	\node [block, below of=feat, text width=20em, yshift=-2em] (neural) {Architecture selection for number of hidden layers in the neural network followed by hyperparameter tuning via grid search. Neural networks with 1, 5, 10, and 15 hidden layers evaluated via 5-fold stratified cross-validation.};
	\node [block, below of=neural, text width=20em, yshift=-1.5em] (annot) {KEGG pathway enrichment of predictive genes selected by the Boruta algorithm.};

>>>>

% Materials and Methods
\section{Methods}
We developed a deep learning model to accurately classify cancer cell lines into therapeutic response groups using data from the Genomics of Drug Sensitivity in Cancer (GDSC) database \cite{gdsc}. Following data collection and curation, we utilized unsupervised clustering algorithms to define two groups based on chemotherapeutic response. Next, we employed a biologically agnostic feature selection algorithm, Boruta, to select statistically relevant genes for our neural network. These transcriptomic features were fed into a neural network that classifies patients into therapeutic response groups as defined by the clustering algorithms. The neural network architecture was optimized using a grid search and evaluated using 5-fold stratified cross-validation. See Fig. \ref{fig:pipeline} for an overview of the data analysis pipeline. The source code can be found on GitHub: \url{https://github.com/tig3r66/youreka_genes}.

% flowchart figure
\begin{figure}[!ht]
	\centering
	\begin{tikzpicture}[node distance=5.5em, auto]
	% blocks
	\node [block, text width=20em] (init) {Obtain data from the GDSC consortium (1,110 cell lines, 251 drugs).};
	\node [block, below of=init, text width=20em, yshift=-1em] (process) {Filter data: retain drugs with $>$80\% data for cell lines (548 cell lines, 117 drugs). Matrix normalization on IC$_{50}$ via z-score.};
	\node [block, below of=process, text width=20em, yshift=-1.5em] (kmeans) {Unsupervised clustering (k-means and principal component analysis) into 2 chemotherapeutic response groups.};
	\node [block, below of=kmeans, text width=20em, yshift=-1.5em] (append) {Appended GDSC Affymetrix Human Genome U219 expression data to the IC$_{50}$ matrix.};
	\node [block, below of=append, text width=20em, yshift=-1.5em] (feat) {Feature selection for relevant genes in RNA microarray data using Boruta algorithm. 300 genes selected at $\alpha=0.05$.};
	\node [block, below of=feat, text width=20em, yshift=-3.25em] (neural) {Architecture selection for number of hidden layers in the neural network followed by hyperparameter tuning via grid search. Neural networks with 1, 5, 10, and 15 hidden layers evaluated via 5-fold stratified cross-validation.};
	\node [block, below of=neural, text width=20em, yshift=-2.75em] (annot) {KEGG pathway enrichment of predictive genes selected by the Boruta algorithm.};

@@
